---
title: "Prob3.10"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "11/11/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let $x_t$ represent the cardiovascular mortality series (cmort) discussed in Example 2.2. 

## (a)
Fit an AR(2) to $x_t$ using linear regression as in Example 3.18.

```{r}
library(astsa) # acf
acf2(cmort, 48)[1:2] # will produce values and a graphic
(regr = ar.ols(cmort, order=2, demean=FALSE, intercept=TRUE)) 
regr$asy.se.coef # standard errors of the estimates
```

The fitted AR(2) model is :

$$ x_t = 11.45 + 0.429 x_{t-1} + 0.442 x_{t-2} + w_t$$

## (b)

Assuming the fitted model in (a) is the true model, find the forecasts over a four-week horizon, $x^n _{n+m}$, for $m = 1, 2, 3, 4$, and the corresponding 95% prediction intervals.

```{r}
n.ahead = 4
sarima.for(cmort, n.ahead, 2, 0, 0) # 4 forecasts with an AR(2) model for cmort
```

Actual values are black. Predictions are red. Confidence intervals are grey (dark=68% and light=95%).
