---
title: "Lab12"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "12/5/2018"
output: pdf_document
---

```{r results='asis', echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(TSA)
library(dplyr)
library(astsa)
library(tseries)
library(forecast)
```

## Fit a seasonal ARIMA model of your choice to the chicken price data in chicken. Check the fit and use the estimated model to forecast the next 12 months.

```{r}
data(chicken)
plot(chicken)
```

# From 2002 to 2016, chicken has an upward trend with some fluctuations.

```{r}
# checking for stationarity
par(mfrow=c(1,2))
acf(chicken, lag.max = 45)
pacf(chicken, lag.max = 45)
```

# acf decays slowly and pacf cuts off immediately. A sign differencing may be required.

```{r}
# Differencing
d.chicken = diff(chicken)
acf2(d.chicken)
```

# After differencing, the acf oscillates with a frequency of 12, pacf cuts off after 12 in the season (s = 12)

```{r}
# Seasonal differencing
sd.chicken = diff(chicken, 12)
acf2(sd.chicken)
```

# After seasonal differencing, acf decays slowly and pacf cuts off after lag 2. Therefore ARIMA (2,1,0) × (1,0,0)12 is suggested.

```{r}
# checking the fit
sarima(chicken, 2,1,0, 1,0,0, 12)
```
 
# Good fit, p-values all lie above the significant level

```{r}
# forcasting
sarima.for(chicken, 12, 2,1,0, 1,0,0, 12)
```

