---
title: "Prob3.42"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "12/5/2018"
output: pdf_document
---

```{r results='asis', echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(TSA)
library(dplyr)
library(astsa)
library(tseries)
library(forecast)
```

## Fit a seasonal ARIMA model of your choice to the unemployment data in UnempRate. Use the estimated model to forecast the next 12 months.

```{r}
data(UnempRate)
plot(UnempRate)
```
 
# Between 1948 and 2017 there are fluctuations in the unemployment rate, 1983 shows the highest peak in unemployment rate.

```{r}
# checking for stationarity
par(mfrow = c(1,2))
acf(UnempRate) 
pacf(UnempRate)
```

## acf decays slowly and pacf cuts off immediately. A sign differencing may be required

```{r}
# Differencing
d.UnempRate = diff(UnempRate) 
acf2(d.UnempRate,48)
```

# Even with the first order differencing there is still residual decay in the acf plots at seasons (s = 12). This suggest a seasonal difference to be applied.

```{r}
# Seasonal differencing
sd.UnempRate = diff(UnempRate,12)
acf2(UnempRate, max.lag = 48)
```

# The pattern in the ACF is not indictive of any simple model as the data is showing abnormalities.

```{r}
auto.arima(UnempRate)
```

# auto.arima suggest ARIMA (3,0,1) × (2,12)12 model

```{r}
# checking the fit
sarima(UnempRate, 3,0,1, 2,1,2, 12)
```
 
# Majority of the p-values are above the significant levels which suggests a good model.

```{r}
# forcasting
sarima.for(UnempRate, 12, 3,0,1, 2,1,2, 12)
```

