---
title: "Lab 11"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "11/19/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Libraries 

```{r}
library(TSA)
library(astsa)
library(datasets)
library(forecast)
library(tseries)
```

#Fit an ARIMA(p, d, q) model to the global temperature data
#globtemp performing all of the necessary diagnostics. After
#deciding on an appropriate model, forecast (with limits) the next
#10 years of the UNdifferenced data. Comment.

```{r}
globtemp
plot(globtemp)
acf(globtemp)
diff_globtemp <- diff(globtemp)
acf2(diff_globtemp)
eacf(diff_globtemp)
```

posibble models: ARMA(1,1) ARMA(0,1)

#ARIMA(1,1,1)
```{r}
sarima(diff_globtemp, 1, 1, 1)
```

Time plot of the residuals does not have any obvious pattern.
The majority of ACF's are all within the limits. The p-values of the
Ljung-Box test are all within the significance level. The QQ plot
affirms normality of the residuals.

#ARIMA(0,1,1)
```{r}
sarima(diff_globtemp, 0, 1, 1)
```
Time plot of the residuals does not have any obvious pattern.
The majority of ACF's are all within the limits. The p-values of the
Ljung-Box test are all within the significance level. The QQ plot
affirms normality of the residuals.

#Compare AIC, AICc, BIC 
AIC, AICc and BIC all prefer ARIMA(1,1,1).

#Final model

$ttable
         Estimate     SE  t.value p.value
ar1       -0.2378 0.0842  -2.8255  0.0055
ma1       -1.0000 0.0190 -52.5815  0.0000
constant   0.0002 0.0002   1.0093  0.3147

$\hat x_t =0.0002 - 0.2378x_{t-1}+ w_t-w_{t-1}$




#forecast
```{r}
n.ahead = 10
sarima.for(globtemp, n.ahead, 1, 1, 1)
```


##Use auto.arima() and compare your results in 1 above.

```{r}
auto.arima(globtemp)
```

auto.arima comfirms the results in 1 above.