---
title: "Lab 10"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "11/13/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(TSA)
library(astsa)
library(datasets)
library(forecast)
library(tseries)
```

Apply the Box-Jenkins method (transform to stationarity if
necessary and identify the time series model(e.g., ARMA(p,q)))
for the luteinizing hormone data. Write down the algebraic
expression of the fitted model. How is your "final" model
compared with an AR(3)?

# plot
```{r}
data(lh)
plot(lh)
```

# estimate p,q
```{R}
acf(lh)
acf2(lh)[1:2]
eacf(lh)

```

candiate models ARMA(1,1) or AR(1)

# check residuals
```{r}
out1 <- arma(x = lh, order = c(1, 0))
summary(out1)
```

```{R}
out2 <- arma(x = lh, order = c(1, 1))
summary(out2)
```

P value of ma(1) is large, which suggests that ar(1) is the better model.

$$ X_t =0.9997 + 0.5861*X_{t-1} + w_t $$

# compare to AR(3)
```{R}
out3 <- arma(x = lh, order = c(3, 0))
summary(out3)
```

Again we see that 2nd and 3rd AR terms are not significant, so AR(1) is prefered over AR(3).

# forecast
Use your final model to forecast the next 24 luteinizing hormone measurements.

```{r}
n.ahead = 24
sarima.for(lh, n.ahead, 1, 0, 0)

```


