---
title: "Lab 6"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "10/10/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(datasets)
library(TSA)
```

## Consider the luteinizing hormone (data(lh) from library(datasets)) in blood samples at 10min intervals from a human female, 48 samples. Is the pattern of the ACF consitent with a stationary AR model?

```{r}
hormone = ts(lh, frequency = 10)
acf(hormone, lag.max = "m")
```

# The pattern is consistent with stationarity as it gradually decays to zero as lag increases

## Simulate an AR(3) using n = 50 and the coefficients = c(0.64, -0.06, -0.22). Using the coefficients, is this a stationary AR(3)? Compare the theoretical ACF with the ACF of the luteinizing hormone data. Observations?

```{r}
ar3 <- arima.sim(list(order = c(3,0,0), ar = c(0.64, -0.06, -0.22)), n = 50)
plot(ar3,ylab=expression(x[t]),xlab="Time",type="l")

z = c(0.64, -0.06, -0.22)
Mod(polyroot(z))

# comparing acf's
acf(ar3,lag.max = "m")
acf(lh, lag.max = "m")
```

# Yes AR(3) is stationary since all its roots has magnitude greater than 1.
# Both data shows stationarity but the luteinizing hormone data decays faster than the AR(3) data.
