---
title: "Lab7"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "10/17/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Simulate an MA(3). Is it invertible? What do you notice about the pattern of the ACF?

```{r}
ma3 = arima.sim(list(order=c(0,0,3), ma= rep(0.9,3)), n=100)
plot(ma3, type = "l", main = "Simulated MA(3) data")
z = c(1,-0.9,-0.9,-0.9)
Mod(polyroot(z)) 
acf(ma3, lag.max = "m", main ="ACF for simulated sample data")
```

# The ma3 model is NOT invertible because only 2 of its roots exceeds 1. All of its roots should exceed 1 for invertibility. The pattern of the ACF as we expect falls off after lag 3 and it is non-stationary as some lags exceed the significant bound.

## 2. Analyze the percentage changes in quarterly personal disposable income for the US, 1970 to 2010. What model can you propose?

```{r}
library(forecast)
library(fpp)
data(usconsumption)
income <- ts(usconsumption[, 2],start=1970:2010,frequency = 4)
ts.plot(income)
acf(income)

d.income <- diff(income)
ts.plot(d.income)
acf(d.income)
```

# Quarterly personal disposable income for the US, 1970 to 2010 is stationary. It can be modeled as a random walk. Changes in quarterly personal disposable income for the US, 1970 to 2010 are non-stationary at lag=1. It can be modeled as a moving average(1).

