---
title: "STAT 5307 Lab 1"
author: "Group4: Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "02SEP2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab 1

## Using RMarkdown, generate n N(0, 1) random data. ( Do this for n = 10, 50, 100, and 1000 )
```{r}
x10 <- rnorm(10, 0, 1)
x50 <- rnorm(50, 0, 1)
x100 <- rnorm(100, 0, 1)
x1000 <- rnorm(1000, 0, 1)

```

## Estimate the first 10 autocovariance and autocorrelations produce plots (trajectory and correlogram). 

### n=10
```{r}
acf(x10, type = "covariance")
```

```{r}
acf(x10, type = "correlation")
```

```{r}
plot( x10
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(x10)))
```

### n=50
```{r}
acf(x50, type = "covariance")
```

```{r}
acf(x50, type = "correlation")
```

```{r}
plot( x50
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(x50)))
```

### n=100
```{r}
acf(x100, type = "covariance")
```

```{r}
acf(x100, type = "correlation")
```

```{r}
plot( x100
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(x100)))
```

### n=1000
```{r}
acf(x1000, type = "covariance")
```

```{r}
acf(x1000, type = "correlation")
```

```{r}
plot( x1000
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(x1000)))
```


## with $\sigma^2 > 1$ :

```{r}
aa <- rnorm(10,0,2)
bb <- rnorm(50,0,2)
cc <- rnorm(100,0,2)
dd <- rnorm(1000,0,2)
```
 
### n=10
```{r}
acf(aa,10)
```
```{r}
plot(aa
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(aa)))
```

### n=50
```{r}
acf(bb,10)
```
```{r}
plot(bb,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(bb)))

```

### n=100
```{r}
acf(cc,10)
```
```{r}
plot(cc
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(cc)))
```

### n=1000
```{r}
acf(dd,10)
```
```{r}
plot(dd
     ,type='l'
     ,ylab="White Noise"
     ,xlab="t"
     ,frame.plot=FALSE
     ,main = paste("n = ",length(dd)))
```


# Compare correlograms. Is there a pattern in the correlograms as the sample size n potentially approaches the population size? If you generate N(0; 2 6= 1) random data, will the ACF change?

As n increases from 10 to 1000, ACF is getting smaller and the significance bounds (confidence limits) become narrower. In all cases, autocorrelation=1 at Lag=0. Values outside the displayed limits have significant autocorrelation. The correlogram does not change when variance is not equal to 1 which indicates the data are truely random.