---
title: "Prob 3.33"
author: "Claudius Taylor, Tom Wilson, Junpu Zhao"
date: "11/19/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries

```{r}
library(TSA)
library(astsa)
library(datasets)
library(forecast)
library(tseries)
```

# fit an ARIMA(p,d,q) model to the global temperature data gtemp performing
# all the necessary diagnostics. After deciding on an appropriate model, 
# forecast(with limits) the next 10 years, comment.

```{r}
gtemp
plot(gtemp)
acf(gtemp)
diff_gtemp <- diff(gtemp)
acf2(diff_gtemp)
eacf(diff_gtemp)
```

ACF cuts off after lag 1, PACF tails off suggesting MA(1) for xt
or ARIMA(0,1,1)
PACF cuts off after lag 3, ACF tails off suggesting AR(3) for xt
or ARIMA(3,1,0)

#ARIMA(0,1,1)
```{r}
sarima(diff_gtemp, 0, 1, 1)
```
Time plot of the residuals does not have any obvious pattern.
The majority of ACF's are all within the limits. The p-values of the
Ljung-Box test are all within the significance level. The QQ plot
affirms normality of the residuals.


#ARIMA(3,1,0)
```{r}
sarima(diff_gtemp, 3, 1, 0)
```
Time plot of the residuals does not have any obvious pattern.
The majority of ACF's are all within the limits. The p-values of the
Ljung-Box test are all within the significance level. The QQ plot
affirms normality of the residuals.

#compare AIC, AICc, BIC
AIC, AICc and BIC all prefer ARIMA(0,1,1).

#forecast
```{r}
n.ahead = 10
sarima.for(gtemp, n.ahead, 0, 1, 1)
```



